Finite-state morphological transducers for three Kypchak languages

### 2014-02-16 kipchak-lrec-2013/kipchak-paper.tex
more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50296 72bbbca6-d526-0410-a7d9-f06f51895060

@@

@@

>>>>

\newcommand{\tags}[1]{\texttt{#1}}

@@

% Russian loans not obeying phonology N5, e.g. ё vs . самолёт

@@

% FIXME: example output

\begin{table*}

% kaz: ^Құдай/Құдай<n><nom>/Құдай<n><attr>/Құдай<n><nom>+е<cop><p3><pl>/Құдай<n><nom>+е<cop><p3><sg>$ ^Өзінің/Өз<prn><ref><p3><sg><gen>$ ^жаратқандарының/жарат<v><tv><ger_past><pl><px3sp><gen>/жарат<v><tv><gpr_past><subst><pl><px3sp><gen>$ ^бәріне/бәрі<prn><qnt><px3sp><dat>$ ^қарап/қара<v><tv><prc_perf>/қара<v><tv><gna_perf>/қара<v><tv><prc_perf>$^,/,<cm>$ ^өте/өт<v><tv><prc_impf>/өте<adv>/өте<v><tv><imp><p2><sg>$ ^жақсы/жақсы<adj>/жақсы<adv>/жақсы<adj><advl>/жақсы<adj><subst><nom>/жақсы<adj>+е<cop><p3><pl>/жақсы<adj>+е<cop><p3><sg>/жақсы<adj><subst><nom>+е<cop><p3><pl>/жақсы<adj><subst><nom>+е<cop><p3><sg>$ ^екенін/е<cop><ger_past><px3sp><acc>$ ^көрді/көр<vaux><ifi><p3><pl>/көр<vaux><ifi><p3><sg>/көр<v><tv><ifi><p3><pl>/көр<v><tv><ifi><p3><sg>$^./.<sent>$ 
% kum: ^Аллагь/аллагь<n><nom>/аллагь<n><attr>$ ^Оьзю/оьз<prn><ref><px3sp><nom>$ ^яратгъан/ярат<v><tv><gpr_past>/ярат<v><tv><gpr_past><subst><nom>/ярат<v><tv><ger_past><nom>/ярат<v><tv><past><p3><sg>$ ^затлагъа/зат<n><pl><dat>$ ^къарап/къара<v><tv><prc_perf>/къара<v><tv><gna_perf>$^,/,<cm>$ ^олар/о<prn><dem><pl><nom>/олар<prn><pers><p3><pl><nom>$ ^бек/бек<adv>$ ^яхшы/яхшы<adj>/яхшы<adj><subst><nom>$ ^экенин/э<cop><ger_past><px3sp><acc>$ ^гёрген/гёр<v><tv><gpr_past>/гёр<v><tv><gpr_past><subst><nom>/гёр<v><tv><ger_past><nom>/гёр<v><tv><past><p3><sg>$^../..<sent>$

\centering

Құдай Өзінің жаратқандарының бәріне қарап, өте жақсы екенін көрді. \\
% ходай Үзенең яратканнарының *бəріне карап, бик яхшы икәнен күрде.
Аллагь Оьзю яратгъан затлагъа къарап, олар бек яхшы экенин гёрген. \\

~\\

\begin{tabular}{lll}
 \textbf{Kazakh} & \textbf{Tatar} & \textbf{Kumyk} \\
\hline
 Құдай\tags{<n><nom>} &  & Аллагь\tags{<n><nom>}  \\
 Өз\tags{<prn><ref><p3><sg><gen>} & & Оьз\tags{<prn><ref><px3sp><nom>} \\
 жарат\tags{<v><tv><ger\_past><pl><px3sp><gen>} & & ярат\tags{<v><tv><gpr\_past>} \\
 бәрі\tags{<prn><qnt><px3sp><dat>} & & зат\tags{<n><pl><dat>} \\
 қара\tags{<v><tv><gna\_perf>} & & къара\tags{<v><tv><gna\_perf>} \\ 
 ,\tags{<cm>} & & ,\tags{<cm>} \\
 ---      & & олар\tags{<prn><pers><p3><pl><nom>} \\
 өт\tags{<adv>} & & бек\tags{<adv>} \\
 жақсы\tags{<adj>} & & яхшы\tags{<adj>} \\
 е\tags{<cop><ger\_past><px3sp><acc>} & & э\tags{<cop><ger\_past><px3sp><acc>} \\
 көр\tags{<v><tv><ifi><p3><sg>} & & гёр\tags{<v><tv><past><p3><sg>} \\  
 .\tags{<sent>} & & .\tags{<sent>} \\
\hline
\end{tabular}

\label{table:exoutput}
\caption{An example of the output of each of the morphological transducers for the same sentence.}

\end{table*}
